\LARGE \bf Predicting Initialization Effectiveness for Trajectory Optimization

### 2014-02-12 paper.tex
acknowledgment
revision

\newcommand{\x}{\mbox{$\mathbf x$}}
\newcommand{\btheta}{\mbox{$\bm \theta$}}
\newcommand{\fsym}{\mbox{$\mathbf z$}}
\newcommand{\fsd}{\mbox{$\mathbf z_{\text{sd}}$}}
\newcommand{\fssd}{\mbox{$\mathbf z_{\text{ssd}}$}}
\newcommand{\fcon}{\mbox{$\mathbf z_{\text{con}}$}}
\newcommand{\fsht}{\mbox{$\mathbf z_{\text{sht}}$}}

@@

\author{Jia Pan \and Pieter Abbeel \thanks{Authors are with the Department of Electrical Engineering and Computer Science, University of California at Berkeley, CA, USA.}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Trajectory optimization is a method for solving motion planning problems by formulating them as non-convex constrained optimization problems. The optimization process, however, can get stuck in local optima that are not collision-free. As a consequence, these methods typically require  multiple initializations, which poses the problem of deciding which initializations to use, given a limited computational budget. In this paper we study the problem of predicting the quality of initializations.  Concretely, we propose a machine learning approach to predicting whether a collision-free solution will be found from a given initialization. We propose a set of trajectory features that encode the obstacle distribution locally around a robot. These features are designed such that they are transferable across different tasks. Experimental evaluation on various planning benchmarks demonstrates the accuracy of our approach when predicting the performance of state-of-the-art trajectory optimization---generalizing over tasks and environments.
\end{abstract}

@@

Despite their success in various applications, trajectory optimization techniques suffer from a critical limitation: their performance heavily depends on the choice of initial trajectories. For instance, certain initializations passing through obstacles in unfavorable ways may get stuck in infeasible solutions and cannot resolve all the collisions in the final outcome, as illustrated in Figure~\ref{fig:failexamples}. As we discuss later, trajectories that pass through the medial axis of an obstacle are prone to getting stuck in a local optimum.

\begin{figure}[t]
\centering
\subfloat[initial path]{\includegraphics[width=0.48\linewidth]{figure/1.png}}
\subfloat[result path]{\includegraphics[width=0.48\linewidth]{figure/2.png}} \\
\subfloat[initial path]{\includegraphics[width=0.48\linewidth]{figure/5.png}}
\subfloat[result path]{\includegraphics[width=0.48\linewidth]{figure/6.png}}
\caption{Failure cases when using the state-of-the-art trajectory optimization method~\cite{Schulman:2013:FLO} in motion planning. (a) shows the initial path for a whole-body planning, which passes through the medial axis of the desk obstacle. (b) is the trajectory optimization outcome, which is stuck in an infeasible condition. (c) shows the initial path for the arm planning and the collision cannot be resolved in the final trajectory shown in (d). }
\label{fig:failexamples}
\end{figure}

We propose a learning-based approach to predicting the quality of a trajectory as an initialization to trajectory optimization based motion planning. Quality is measured by whether the optimization converges to a collision free trajectory connecting start and goal. 
For the learning algorithm, we first design a set of trajectory features, which are relevant to a trajectory's effectiveness as an initialization. These features are both task-independent and scene-independent, and hence they are transferable among different planning tasks. We use three types of features: first, spatial signed distance (SSD) vectors for each robot link; second, the difference between SSD vectors for the same robot link in adjacent trajectory waypoints and for adjacent robot links in the same waypoint; and third, the spherical harmonics transform of SSD vectors. These features are able to encode the obstacle distribution locally around a robot trajectory and thus are essential for predicting a trajectory's behavior under a trajectory optimization algorithm. Next, we calculate the features for a set of randomly selected trajectories for different tasks in various benchmarks. To generate a labeled dataset, we run trajectory optimization for a large number of tasks and environments. Based on the dataset, we can learn a prediction function for trajectory effectiveness, which can be integrated with trajectory optimization algorithms to improve their solving performance and solution quality.

The rest of the paper is organized as follows. We survey related work on trajectory optimization and initial value selection for general optimization problems in Section~\ref{sec:related}. In Section~\ref{sec:probstatement}, we define the problem of predicting the quality of a trajectory as an initial guess. In Section~\ref{sec:predict}, we present the details about how to implement such prediction function using machine learning and a set of carefully selected features. Section~\ref{sec:experiment} describes the results of our method on various planning benchmarks.

\section{Related Work}
\label{sec:related}
Trajectory optimization is an established method in robotics to generate a high quality path from an initial trajectory that may be in-collision or dynamically infeasible; it works by solving an optimization problem on trajectory space. Many approaches use a spline-based or waypoint-based representation for the trajectory and use cost gradient information for minimization, including CHOMP~\cite{Ratliff:2009:CGO}, STOMP~\cite{Kalakrishnan:2011:STOMP} and TrajOpt~\cite{Schulman:2013:FLO}. There is extensive, closely-related work in optimal control. The optimal control literature focuses less on collisions and more on systems with complicated dynamical properties. Some of the most notable work includes DDP (Differential Dynamic Programming)~\cite{Jacobson:1970:DDP,Atkeson:1994:ULT}, Iterative LQR (iLQR)~\cite{Todorvo:2005:iLQG}, Approximate Inference Control~\cite{Toussaint:2009:RTO} and optimization-based control that involves contacts~\cite{Mordatch:2012:DCB,Tassa:2012:SSC,Erez:2012:TOD,Posa:2013:DTO}.

One main challenge of trajectory optimization is that it is sensitive to the choice of initial trajectories and may get stuck in infeasible local optima. This is because the set of all collision-free configurations is highly non-convex, so the optimization is prone to getting stuck in local minima.

For general non-convex optimization problems, the dependence on initialization is also a well-known challenge. A simple and popular solution is using multiple random initializations to help the algorithms escape from bad local optima. 
Recent work~\cite{Cassioli:2012:MLG} uses machine learning to learn the relationship between the starting point of an optimization algorithm and the objective value of the final outcome and is closely related with our method. The main difference is that our approach is for trajectory optimization, where generating a feasible trajectory is more important than decreasing the absolute cost of the objective function. Moreover, trajectory optimization can exploit workspace or physics based heuristics for selecting good initial guesses, which are not available in general optimization problems.

Another closely related line of work is on trajectory prediction~\cite{Jetchev:2013:FMP}, where a high-dimensional feature vector is used to capture information about the proximity of a robot to the obstacles. After a sparse regularized feature selection process, machine learning algorithms are used to predict a good path in a new environment from the database of previous trajectories and environments. The predicted result may not be feasible and hence requires another trajectory optimization step to resolve all the constraint violations. This post-processing might fail as the predicted trajectory may be a bad initialization for trajectory optimization.

Prior work has also considered designing trajectory features for various applications. For example, Bentivegna et al.~\cite{Bentivegna:2006:LST} and Stolle et al.~\cite{Stolle:2007:TPT} represented a library of trajectories by their feature vectors in order to make control policy transfer more convenient. 
Berenson et al.~\cite{Berenson:2012:RPP} designed several criteria and features to evaluate 
whether a trajectory from a path library is likely to be reused in a new planning scene after suitable repair operations. Trajectory features can also be used to predict the class labels of moving objects based on their trajectories~\cite{Lee:2008:TTC}. Moreover, features have been designed to compress trajectory libraries~\cite{Arikan:2006:CMC}. However, all these features only capture properties of the trajectory itself. In contrast, the features proposed in our paper encode the interaction between the trajectory and the surrounding environment, which is critical for predicting how fast a trajectory will be pushed away from the obstacles by a trajectory optimization algorithm.

@@

Robotic motion planning problems can be formulated as non-convex optimization problems, i.e., minimize an objective subject to inequality and equality constraints:

\begin{equation*}
\begin{aligned}
& \underset{\x}{\text{minimize}}
& & f(\x) \\
& \text{subject to}
& & g_i(\x) \leq 0, \; i = 1, 2, \dots, n_{\text{ieq}} \\
&&& h_i(\x) = 0, \; i = 1, 2, \dots, n_{\text{eq}}
\end{aligned}
\end{equation*}
where $f$, $g_i$ and $h_i$ are scalar functions. For planning problems that involve only kinematics and represent the trajectory as a sequence of $T$ waypoints, the optimization variable $\x$ is of the form of $\x = \btheta_{1:T}$, where $\btheta_t \in \mathbb{R}^K$ denotes the configuration at the $t$-th waypoint for a system with $K$ degrees of freedom. For problems with dynamics, the optimization variable $\x$ may also include velocities $\dot{\btheta_t}$ and torques ${\bm \tau}_t$.

Collision avoidance is one of the most important inequality constraints in motion planning. But it is difficult to formulate in closed form and hence is challenging for optimization. Various approximations for collision avoidance constraints have been proposed, including distance fields~\cite{Khatib:1985:ROA, Ratliff:2009:CGO}, penetration depth~\cite{Cameron:1997:CMP}, and swept volume~\cite{Schulman:2013:FLO}. Besides collision avoidance, common inequality constraints include kinematic constraints (e.g., joint limits and static stability constraints), kinodynamic constraints (e.g., bounded velocity, acceleration or force/torque) or dynamic constraints (e.g., dynamic stability constraints). One example of equality constraints is the end-effector pose constraint where the robot must reach a target pose at the end of the trajectory.

@@

Trajectory optimization is a challenging non-convex problem, and many approaches have been presented to solve it efficiently and effectively. Among them are sequential convex optimization~\cite{Schulman:2013:FLO}, covariant Hamiltonian optimization~\cite{Ratliff:2009:CGO}, and stochastic optimization~\cite{Kalakrishnan:2011:STOMP}. However, given an initial trajectory that contains collisions, these methods are not guaranteed to find a collision-free solution as the collision avoidance terms in the optimization are non-convex. Figure~\ref{fig:badgradient} shows some scenarios illustrating how trajectory optimization tends to get stuck in local optima that are not collision-free.

\begin{figure}[t]
\centering
\subfloat[]{\includegraphics[width=0.42\linewidth]{figure/medial_axis.pdf}
}
\subfloat[]{\includegraphics[width=0.55\linewidth]{figure/distance_field.pdf}} \\
\subfloat[]{\includegraphics[width=0.48\linewidth]{figure/multiple_obstacle.pdf}}
\subfloat[]{\includegraphics[width=0.48\linewidth]{figure/multiple_link.pdf}}
\caption{Illustration of typical reasons for trajectory optimization to get stuck in local optima that are not collision-free. (a) The gradient based on penetration depth may push waypoints in in-consistent directions. (b) The gradient based on distance fields has the same problem. (c) When a robot collides simultaneously with multiple obstacles, the robot may get stuck in an infeasible local optimum as different obstacles push the robot in different directions. (d) For a robot with multiple links, the gradient may result in in-consistent directions for different links. $\btheta_i$ in these figures denote different trajectory waypoints.}
\label{fig:badgradient}
\end{figure}

@@

\subsection{Outline of Proposed Approach}
We first design trajectory features that are potentially useful to distinguish good and bad initializations, which is based on the assumption that trajectories converging to feasible solutions share common properties. The features extracted from a trajectory $\btheta_{1:T}$ are denoted as a vector $\fsym = E(\btheta_{1:T})$, where $E(\cdot)$ is the extraction function. For each trajectory, we use a binary variable $y$ to denote whether the trajectory optimization succeeds ($y = 1$) or fails ($y = 0$).

Given a set of $N$ trajectories, we can build a training set $\{\fsym_i, y_i\}_{i=1}^N$. Based on the data set, we can learn a classifier $C(\fsym) = \{0, 1\}$, where $1$ means a feasible solution would be returned and the corresponding trajectory is potentially a good initial guess. The classifier $C(\cdot)$ is the predicator for a trajectory's effectiveness.

@@

In this section, we discuss how to design trajectory features to distinguish good and bad initializations and how to use the extracted features to predict the effectiveness of a new trajectory.

Before extracting features from a trajectory, we first perform re-sampling on it so that the waypoints are uniformly distributed between two end-points. We perform this pre-processing based on what was pointed out by Ratliff et al.~\cite{Ratliff:2009:CGO}: the objective function should be invariant to the reparameterization of the trajectory so that the gradient will pull a trajectory out of obstacles instead of encouraging it to quickly jump through obstacles. Some recent work such as~\cite{Schulman:2013:FLO} handles the `jumping through problem' in a completely different way, but in practice, it is observed that a re-sampling pre-processing can still result in better convergence. Our re-sampling process excludes the difference of trajectories in parameterization and hence makes it easier to compare them. After re-sampling, each trajectory has $T$ waypoints. This also guarantees our feature extraction gives fixed-length vectors for different trajectories.

\subsection{Background: Signed Distances}
We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Formally, the signed distance is defined between two sets $\mathcal A$ and $\mathcal B$ in workspace as follows:
\begin{align}
\text{sd}(\mathcal A, \mathcal B) = \text{dist}(\mathcal A, \mathcal B) - \text{penetration}(\mathcal A, \mathcal B).
\end{align}
Here $\text{dist}(\mathcal A, \mathcal B)$ is the distance between $\mathcal A$ and $\mathcal B$, i.e., the length of the smallest relative translation $\mathbf T$ that puts these two non-overlapped sets in contact:
\begin{align}
\text{dist}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B \neq \emptyset\}.
\end{align}
And $\text{penetration}(\mathcal A, \mathcal B)$ is the (translational) penetration depth between $\mathcal A$ and $\mathcal B$, i.e., the minimum translation $\mathbf T$ that makes two overlapped sets out of contact:
\begin{align}
\text{penetration}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B = \emptyset\}.
\end{align}
The signed distance computation can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}. Signed distance also provides a normal vector $\mathbf n = \frac{\mathbf T}{\|\mathbf T\|}$, i.e., the direction that two objects can be most efficiently pushed away from each other. Such a direction is important for a trajectory's effectiveness: As we showed in Figure~\ref{fig:badgradient}, a bad initialization tends to have adjacent waypoints with very different pushing away directions.

\subsection{Feature I: Spatial Signed Distance Vectors}
Based on the signed distances formulation, a natural choice for trajectory feature design is to collect the signed distances and normals between all pairs of robot links and environment obstacles, for each waypoint in the trajectory. More formally, given a robot with $n$ links $\mathcal A_i$ in an environment with $m$ obstacles $\mathcal B_j$, the feature vector would be $\fsd = \{\text{sd}_{i,j,t}, \mathbf n_{i,j,t}\}_{1\leq i \leq n, 1\leq j \leq m, 1\leq t \leq T}$, where $\text{sd}_{i,j,t} = \text{sd}(\mathcal A_i(t), \mathcal B_j)$ is the signed distance value between $\mathcal A_i$ and $\mathcal B_j$ at waypoint $\btheta_t$ and $\mathbf n_{i,j,t}$ is the corresponding normal. This feature vector has a dimension of $4\times m \times n \times T$ and indeed provides a complete description for the obstacle distribution around the trajectory, which is critical when predicting trajectory's effectiveness. However, the feature dimension will depend on $m$, the number of obstacles in the environment, which makes this feature not transferable between different environments. Moreover, in many real-world systems, there may be a large number of obstacles (e.g., the obstacles from sensor data are usually represented as thousands of boxes~\cite{Sucan:2010:CPT}). This will result in a long feature vector and brings challenges for both data storage and the following learning procedure. As a result, we need to summarize this feature in a more compact way but keep its rich information as much as possible.

\begin{figure}[t]
\centering
\subfloat[]{\includegraphics[width=0.52\linewidth]{figure/penetration_encode.pdf}}
\subfloat[]{\includegraphics[width=0.47\linewidth]{figure/discrete_directions.pdf}}
\caption{Spatial signed distance vectors: (a) shows a 2D example about spatial signed distance vectors generation. The surrounding space around a robot link (the cyan color part) is uniformly divided by a spatial vector dictionary with $12$ prototype directions $\{\mathbf b_1, ..., \mathbf b_{12}\}$. The signed distance value $\text{sd}_j$ between the robot and the $j$-th obstacle (the grey parts) is then accumulated toward a prototype direction $\mathbf b_k$ which has the smallest angle with the corresponding signed distance normal $\mathbf n_j$, i.e., with the smallest $|\mathbf n_j \cdot \mathbf b_k|$. (b) In 3D case, we use the approach introduced in~\cite{Yershova:2010:GUI} to generate dictionary vectors that can make equisurface partition on a sphere surface. }
\label{fig:trajectoryfeaturePD}
\end{figure}

>>>>

\newcommand{\x}{\mbox{$\mathbf x$}}
\newcommand{\btheta}{\mbox{$\bm \theta$}}
\newcommand{\fsym}{\mbox{$\mathbf z$}}
\newcommand{\fsd}{\mbox{$\mathbf z_{\text{sd}}$}}
\newcommand{\fssd}{\mbox{$\mathbf z_{\text{ssd}}$}}
\newcommand{\fcon}{\mbox{$\mathbf z_{\text{con}}$}}
\newcommand{\fsht}{\mbox{$\mathbf z_{\text{sht}}$}}
\newcommand{\fconv}{\mbox{$\mathbf z_{\text{conv}}$}}

@@

\author{Jia Pan \and Zhuo Chen \and Pieter Abbeel \thanks{Authors are with the Department of Electrical Engineering and Computer Science, University of California at Berkeley, CA, USA.}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Trajectory optimization is a method for solving motion planning problems by formulating them as non-convex constrained optimization problems. The optimization process, however, can get stuck in local optima that are in collision. As a consequence, these methods typically require multiple initializations; this poses the problem of deciding which initializations to use, given a limited computational budget. In this paper we study the problem of predicting the quality of initial trajectories. Concretely, we propose a machine learning approach to predict whether a collision-free solution will be found from a given initialization. We present a set of trajectory features that encode the obstacle distribution locally around a robot. These features are designed such that they are transferable across different tasks. Experimental evaluation on various planning benchmarks demonstrates the accuracy of our approach when predicting the performance of state-of-the-art trajectory optimization---generalizing over tasks and environments.
\end{abstract}

@@

Despite their success in various applications, trajectory optimization techniques suffer from a critical limitation: their performance heavily depends on the choice of initial trajectories. For instance, certain initializations passing through obstacles in unfavorable ways may get stuck in infeasible solutions and cannot resolve all collisions in the final outcome, as illustrated in Figure~\ref{fig:failexamples}. As we discuss later, trajectories that pass through the medial axis of an obstacle are prone to getting stuck in a local optimum.

\begin{figure}[!h]
\centering
\subfloat[initial path]{\includegraphics[width=0.38\linewidth]{figure/1.png}} \qquad
\subfloat[result path]{\includegraphics[width=0.38\linewidth]{figure/2.png}} \\
\subfloat[initial path]{\includegraphics[width=0.38\linewidth]{figure/5.pdf}} \qquad
\subfloat[result path]{\includegraphics[width=0.38\linewidth]{figure/6.pdf}}
\caption{Failure cases when using the state-of-the-art trajectory optimization method~\cite{Schulman:2013:FLO} for motion planning. (a) shows the initial path for a whole-body planning, which passes through the medial axis of the desk obstacle. (b) is the trajectory optimization outcome, which is stuck in an infeasible condition and the trajectory is pulled apart by virtue of being on both sides of an obstacle. (c) shows the initial path for the arm planning and the collision cannot be resolved in the final trajectory shown in (d). The main difference between trajectories in (c) and (d) is marked by dashed red boxes.}
\label{fig:failexamples}
\end{figure}

We propose a learning-based approach to predict the quality of a trajectory as an initialization to the optimization-based motion planning. Here, a trajectory's quality is measured by whether the optimization problem converges to a collision-free trajectory when using the given trajectory as the initial guess. In other words, the quality is defined in a qualitative manner, i.e. `good' or `bad'.
For the learning algorithm, we first design a set of trajectory features, which are relevant to a trajectory's effectiveness as an initialization. These features are both task-independent and scene-independent, and hence they are transferable among different planning tasks. We use four types of features: 1) spatial signed distance (SSD) vectors for each robot link; 2) difference between SSD vectors for the same robot link in adjacent trajectory waypoints and for adjacent robot links in the same waypoint; 3) spherical harmonics transform of SSD vectors; and 4) features measuring the non-convexity of the constrained optimization problem close to the trajectory. These features encode the obstacle distribution locally around a robot trajectory and thus are essential for predicting a trajectory's behavior under a trajectory optimization algorithm. These features are calculated for a set of randomly selected trajectories for different tasks in various benchmarks. Next, we run trajectory optimization using these trajectories as initial guesses and obtain a labeled dataset. In the dataset, each trajectory is augmented with a `good' or `bad' label according to whether the optimization converges to a collision-free trajectory or not. With this dataset as the training set, we learn a predict function for trajectory effectiveness. We evaluate the accuracy of the learned predict function by counting the ratio of correct predictions for a set of new trajectories on various benchmarks.

\section{Related Work}
\label{sec:related}
Trajectory optimization is an established method in robotics to generate a high quality path from an initial trajectory that may be in-collision or dynamically infeasible. It has also been used as a post-processing phase to remove redundant or jerky motion in trajectories generated by traditional planning algorithms~\cite{Latombe:1991:RMP,Lavalle:2006:PA}. Trajectory optimization works by solving an optimization problem on trajectory space. Many approaches use a spline-based or waypoint-based representation for trajectory and use cost gradient information for minimization, including CHOMP~\cite{Ratliff:2009:CGO}, STOMP~\cite{Kalakrishnan:2011:STOMP} and TrajOpt~\cite{Schulman:2013:FLO}. There is extensive, closely-related work in optimal control, which focuses less on collisions and more on systems with complicated dynamical properties. Some of the most notable work includes differential dynamic programming~\cite{Jacobson:1970:DDP,Atkeson:1994:ULT}, iterative LQR~\cite{Todorvo:2005:iLQG}, approximate inference control~\cite{Toussaint:2009:RTO} and optimization-based control involving contacts~\cite{Mordatch:2012:DCB,Tassa:2012:SSC,Erez:2012:TOD,Posa:2013:DTO}.

One main challenge of trajectory optimization is its sensitivity to the choice of initial guesses and may get stuck in infeasible local optima, since the collision-free constraint is highly non-convex.

For general non-convex optimization problems, the dependence on initialization is also a well-known challenge. A simple and popular solution is using multiple random initializations to help the algorithms escape from bad local optima. 
One recent work closely related with our work is~\cite{Cassioli:2012:MLG}, which uses machine learning to build the relationship between the starting point of an optimization algorithm and the objective value of the final outcome. The main difference is that our approach is for trajectory optimization, where generating a feasible trajectory is more important than decreasing the absolute cost of the objective function. Moreover, trajectory optimization can exploit workspace heuristics for selecting good initial guesses, and such heuristics are not available in general optimization problems. Another closely related line of work is the trajectory prediction~\cite{Jetchev:2013:FMP}, where learning algorithms are used to predict a good path in a new environment from a database of demonstrated trajectories. The predicted result may not be feasible and hence requires another trajectory optimization step to resolve all violations.

Prior work has considered designing trajectory features for various applications. For example, Bentivegna et al.~\cite{Bentivegna:2006:LST} and Stolle et al.~\cite{Stolle:2007:TPT} represented a library of trajectories by feature vectors for control policy transferring. 
Berenson et al.~\cite{Berenson:2012:RPP} designed several criteria and features to evaluate whether a trajectory from a path library is likely to be reused in a new planning scene after suitable repair operations. Trajectory features are also used to predict class labels of moving objects based on their trajectories~\cite{Lee:2008:TTC}. Moreover, features have been designed to compress trajectory libraries~\cite{Arikan:2006:CMC}. However, all these features only capture properties of a trajectory itself. In contrast, the features proposed in our paper encode the interaction between the trajectory and the surrounding environment, which is critical for predicting how fast a trajectory can be pushed away from the obstacles by a trajectory optimization algorithm.

@@

Robotic motion planning problems can be formulated as non-convex optimization problems, i.e., minimize an objective subject to inequality and equality constraints:
\begin{align}
\label{eq:opt}
& \underset{\x}{\text{minimize}}
& & f(\x) \\
& \text{subject to}
& & g_i(\x) \leq 0, \; i = 1, 2, \dots, n_{\text{ieq}} \notag \\
&&& h_i(\x) = 0, \; i = 1, 2, \dots, n_{\text{eq}} \notag
\end{align}
where $f$, $g_i$ and $h_i$ are scalar functions. For planning problems that involve only kinematics and represent the trajectory as a sequence of $T$ waypoints, the optimization variable $\x$ is of the form of $\x = \btheta_{1:T}$, where $\btheta_t \in \mathbb{R}^K$ denotes the configuration at the $t$-th waypoint for a system with $K$ degrees of freedom. For problems with dynamics, the optimization variable $\x$ may also include velocities $\dot{\btheta_t}$ and torques ${\bm \tau}_t$.

\begin{figure}[!h]
\centering
\subfloat[]{\includegraphics[width=0.34\linewidth]{figure/medial_axis.pdf}} \qquad
\subfloat[]{\includegraphics[width=0.44\linewidth]{figure/distance_field.pdf}} \\
\subfloat[]{\includegraphics[width=0.38\linewidth]{figure/multiple_obstacle.pdf}} \qquad
\subfloat[]{\includegraphics[width=0.38\linewidth]{figure/multiple_link.pdf}}
\caption{Illustration of typical reasons for trajectory optimization to get stuck in local optima that are not collision-free. The gradient based on penetration depth (a) or distance fields (b) may push waypoints in in-consistent directions. When a robot collides simultaneously with multiple obstacles (c), the robot may get stuck in an infeasible local optimum since different obstacles push the robot in different directions. For a robot with multiple links (d), the gradient may result in in-consistent directions for different links. $\mathbf x_i$ in these figures denote different trajectory waypoints.}
\label{fig:badgradient}
\end{figure}

Collision avoidance is one of the most important inequality constraints in motion planning. But it is difficult to be formulated in a closed form and hence is challenging for optimization. Various approximations for collision avoidance constraints have been proposed, including distance fields~\cite{Khatib:1985:ROA, Ratliff:2009:CGO}, penetration depth~\cite{Cameron:1997:CMP}, and swept volume~\cite{Schulman:2013:FLO}. Besides collision avoidance, common inequality constraints include kinematic constraints (e.g., joint limits), kinodynamic constraints (e.g., bounded velocity or acceleration) or dynamic constraints (e.g., dynamic stability constraints). One example of equality constraints is the end-effector pose constraint where the robot must reach a target pose.

@@

Trajectory optimization is a challenging non-convex problem, and many approaches have been presented to solve it effectively. Among them are sequential convex optimization~\cite{Schulman:2013:FLO}, covariant Hamiltonian optimization~\cite{Ratliff:2009:CGO}, and stochastic optimization~\cite{Kalakrishnan:2011:STOMP}. However, given an initial trajectory that contains collisions, these methods are not guaranteed to find a collision-free solution due to the non-convex constraints in the optimization. Figure~\ref{fig:badgradient} shows some scenarios illustrating how trajectory optimization tends to get stuck in local optima that are not collision-free.

@@

\subsection{Outline of Proposed Approach}
The problem to predict whether a given trajectory is a good or bad initial guess for a trajectory optimization algorithm can be formalized as a two-class classification problem. In fact, a trajectory is good if and only if it locates in the basin of attraction of a local optima that is collision-free, as shown in Figure~\ref{fig:overview}. In this way, all the trajectories in the trajectory space can be categorized into two classes, according to whether they are attracted to a collision-free or an in-collision local optimum.

\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{figure/overview.pdf}
\caption{Trajectory quality prediction as a two-class classification problem. The curve is the trajectory cost manifold: the blue part corresponds to the collision-free trajectories while the black parts correspond to in-collision ones. The trajectory space can be classified into two classes (red and green) according to the basin of attraction that each trajectory locates in. }
\label{fig:overview}
\end{figure}

For the classification problem, we first design trajectory features that are potentially useful to distinguish good and bad initializations. The features extracted from a trajectory $\btheta_{1:T}$ are denoted as a vector $\fsym = E(\btheta_{1:T})$, where $E(\cdot)$ is the extraction function. For each trajectory, we use a binary variable $y$ to denote whether the trajectory optimization succeeds ($y = 1$, a good initial guess) or fails ($y = 0$, a bad initial guess).

Given a set of $N$ trajectories, we build a training set $\{\fsym_i, y_i\}_{i=1}^N$. Based on the data set, we learn a classifier $C(\fsym) = \{0, 1\}$, which is the predicator for a trajectory's effectiveness. Given a new trajectory as input, $C(\cdot)$ will provide its qualitative evaluation (good or bad) as output. The computed predicator $C(\cdot)$ may change among different planning methods, because different optimization approaches may label trajectories in the training set differently. However, our learning method is general and works for all trajectory optimization based planning methods.

@@

In this section, we design trajectory features to distinguish good and bad initializations and use the extracted features for predicting the effectiveness of a new trajectory.

Before extracting features from a trajectory, we first perform re-sampling on it so that waypoints are uniformly distributed between two end-points.
%We perform this pre-processing based on what was pointed out by Ratliff et al.~\cite{Ratliff:2009:CGO}: the objective function should be invariant to the re-parameterization of the trajectory so that the gradient will pull a trajectory out of obstacles instead of encouraging it to quickly jump through obstacles. Some recent work such as~\cite{Schulman:2013:FLO} handles the `jumping through problem' in a completely different way, but in practice, it is observed that a re-sampling pre-processing can still result in better convergence. 
The re-sampling process excludes the difference of trajectories in parameterization and hence makes it easier to compare them. After re-sampling, each trajectory has $T$ waypoints. This also guarantees our feature extraction gives fixed-length vectors for different trajectories.

%\subsection{Background: Signed Distances}
%We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Formally, the signed distance is defined between two sets $\mathcal A$ and $\mathcal B$ in workspace as follows:
%\begin{align}
%\text{sd}(\mathcal A, \mathcal B) = \text{dist}(\mathcal A, \mathcal B) - \text{penetration}(\mathcal A, \mathcal B).
%\end{align}
%Here $\text{dist}(\mathcal A, \mathcal B)$ is the distance between $\mathcal A$ and $\mathcal B$, i.e., the length of the smallest relative translation $\mathbf T$ that puts these two non-overlapped sets in contact:
%\begin{align}
%\text{dist}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B \neq \emptyset\}.
%\end{align}
%And $\text{penetration}(\mathcal A, \mathcal B)$ is the (translational) penetration depth between $\mathcal A$ and $\mathcal B$, i.e., the minimum translation $\mathbf T$ that makes two overlapped sets out of contact:
%\begin{align}
%\text{penetration}(\mathcal A, \mathcal B) = \inf\{\|\mathbf T\| \mid (\mathbf T + \mathcal A) \cap \mathcal B = \emptyset\}.
%\end{align}
%The signed distance computation can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}. Signed distance also provides a normal vector $\mathbf n = \frac{\mathbf T}{\|\mathbf T\|}$, i.e., the direction that two objects can be most efficiently pushed away from each other. Such a direction is important for a trajectory's effectiveness: As we showed in Figure~\ref{fig:badgradient}, a bad initialization tends to have adjacent waypoints with very different pushing away directions.

\subsection{Background: Signed Distances}
We design features to encode the obstacle distribution locally around a trajectory, as such information is closely related with a trajectory's effectiveness. Intuitively speaking, if a trajectory is deep inside the obstacles, it is less likely to be pushed out of them; whereas if it is mostly outside the obstacles, it should converge to a collision-free solution. To quantify the obstacle distribution surrounding a trajectory, we use signed distances and normals, which are used in previous trajectory optimization algorithms such as~\cite{Schulman:2013:FLO}. Informally, signed distance between two objects is the length of the smallest translation that puts them in contact; signed normal is the direction of such translation. The signed distance computation can be efficiently computed for convex objects using GJK algorithm~\cite{Gilbert:1988:GJK} and EPA algorithm~\cite{Bergen:2001:EPA}.

\subsection{Feature I: Spatial Signed Distance Vectors}
Based on the signed distances formulation, a natural choice for designing trajectory feature is to collect signed distances and normals between all pairs of robot links and environment obstacles, for each waypoint in the trajectory. More formally, given a robot with $n$ links $\mathcal A_i$ in an environment with $m$ obstacles $\mathcal B_j$, the feature vector would be $\fsd = \{\text{sd}_{i,j,t}, \mathbf n_{i,j,t}\}_{1\leq i \leq n, 1\leq j \leq m, 1\leq t \leq T}$, where $\text{sd}_{i,j,t} = \text{sd}(\mathcal A_i(t), \mathcal B_j)$ is the signed distance value between $\mathcal A_i$ and $\mathcal B_j$ at waypoint $\btheta_t$, and $\mathbf n_{i,j,t}$ is the corresponding normal. This feature vector is of dimension $4\times m \times n \times T$ and provides a complete description for the obstacle distribution around the trajectory. However, the dimension of this feature depends on $m$, the number of obstacles in the environment, which makes this feature not transferable among different environments. Moreover, in many real-world systems, there may be a large number of obstacles (e.g., the obstacles from sensor data are usually represented as thousands of boxes~\cite{Sucan:2010:CPT}). This would result in a long feature vector and brings challenges for both data storage and the following learning procedure. Thus we choose to summarize this feature in a more compact manner but keep its rich information as much as possible.

\begin{figure}[!h]
\centering
\subfloat[]{\includegraphics[width=0.52\linewidth]{figure/penetration_encode.pdf}}
\subfloat[]{\includegraphics[width=0.47\linewidth]{figure/discrete_directions.pdf}}
\caption{Spatial signed distance vectors: (a) shows a 2D example about SSD vector generation. The surrounding space around a robot link (the cyan color part) is uniformly divided by a spatial vector dictionary with $12$ directions $\{\mathbf b_1, ..., \mathbf b_{12}\}$. The signed distance value $\text{sd}_j$ between the robot and the $j$-th obstacle (the grey part) is then accumulated toward a direction $\mathbf b_k$, which has the smallest angle with the corresponding signed distance normal $\mathbf n_j$, i.e., with the smallest $|\mathbf n_j \cdot \mathbf b_k|$. (b) In 3D case, we use the approach introduced in~\cite{Yershova:2010:GUI} to generate dictionary vectors that make equisurface partition on a sphere surface. }
\label{fig:trajectoryfeaturePD}
\end{figure}
